### 2014-06-11 Studies/2012_Chicago/Report/Chicago_Fire.tex
Chicago_Fire: update report and bib file

Prior to door failure, the elevated temperatures on the second floor were limited to the rear porch, Fig.~(\ref{fig:temp_top_159}). Temperatures are highest at the Beta-Charlie corner where there is sufficient oxygen due to self-ventilation. Five seconds after the door fails, hot gases have traveled the length of the interior hallway and approached the doorway into the kitchen, approximately 4.5 m (14.8 ft) away from the source fire (cf Fig.~(\ref{fig:temp_top_165})). One minute after the door fails, the hallway temperatures are nominally at or above 260 $^{\circ}$C (500 $^{\circ}$F) through out the hallways corridor (cf Fig.~(\ref{fig:temp_top_220})). This illustrates the the high temperature gases seen five seconds after failure were not just a burst but rather the front of a sustained exhaust gas flow from the fire burning on the rear porch. The conditions in this hallway changed from tenable to high-hazard very rapidly following the door failure. Fig.~(\ref{fig:temp_top_220}) also shows that that temperatures in the kitchen dropped below 100 $^{\circ}$C (212 $^{\circ}$F) as the large supply of air cooled/diluted the exhaust gases. Based on post-incident images of the kitchen and second floor stairwell, these predicted temperatures are realistic.

\begin{figure}[h!]
\centering
\includegraphics[width=.7\textwidth]{../Figures/west_50th_baseline_top_159}
\caption{PLACEHOLDER: Temperature contours on the second floor, 1 second prior to door failure, 6ft above the floor.}
\label{fig:temp_top_159}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=.7\textwidth]{../Figures/west_50th_baseline_top_165}
\caption{PLACEHOLDER: Temperature contour on the second floor, 5 seconds after door failure, 6ft above the floor.}
\label{fig:temp_top_165}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=.7\textwidth]{../Figures/west_50th_baseline_top_200}
\caption{PLACEHOLDER: Temperature contour on the second floor, 60 seconds after door failure, 6ft above the floor.}
\label{fig:temp_top_220}
\end{figure}

\section{Tactical Considerations}
In the last fifteen years, there have been fires throughout the country where the change in a flow path has had a significant impact on firefighter and occupant safety:

{\bf UNDER DEVELOPMENT}

\begin{itemize}
\item NE Washington D.C., 1999
\item Keokuk, IA, 1999
\item Houston, TX, 2011
\item Riverdale Heights, MD, 2012 
\end{itemize}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=.80\textwidth]{../Figures/50th_Place_1st_Floor}
\caption {Dimensioned drawing of 1$^{st}$ floor}
\label{fig:first_floor}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=.80\textwidth]{../Figures/50th_Place_2nd_Floor}
\caption {Dimensioned drawing of 2$^{nd}$ floor}
\label{fig:second_floor}
\end{figure}

>>>>

Prior to door failure, the elevated temperatures on the second floor were limited to the rear porch, Fig.~(\ref{fig:temp_top_159}). Temperatures are highest at the Beta-Charlie corner where there is sufficient oxygen due to self-ventilation. Five seconds after the door fails, hot gases have traveled the length of the interior hallway and approached the doorway into the kitchen, approximately 4.5 m (14.8 ft) away from the source fire (cf Fig.~(\ref{fig:temp_top_165})). One minute after the door fails, the hallway temperatures are nominally at or above 260 $^{\circ}$C (500 $^{\circ}$F) through out the hallways corridor (cf Fig.~(\ref{fig:temp_top_220})). This illustrates the the high temperature gases seen five seconds after failure were not just a burst but rather the front of a sustained exhaust gas flow from the fire burning on the rear porch. The conditions in this hallway changed from tenable to high-hazard very rapidly following the door failure. Fig.~(\ref{fig:temp_top_220}) also shows that the temperatures in the kitchen were at or below 90 $^{\circ}$C (194 $^{\circ}$F) and temperatures at the top of the stairs were at or below 70 $^{\circ}$C (158 $^{\circ}$F). At these points in the structure, the exhaust gases in the flow path are air cooled/diluted by the inflow ambient gas. Based on post-incident images of the kitchen and second floor stairwell, these predicted temperatures are on the appropriate order or magnitude.

\begin{figure}[h!]
\centering
\includegraphics[width=.7\textwidth]{../Figures/west_50th_baseline_top_159_6ft}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{Temperature contours on the second floor, 1 second prior to door failure, 1.83 m (6 ft) above the floor.}
\label{fig:temp_top_159}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=.7\textwidth]{../Figures/west_50th_baseline_top_165_6ft}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{Temperature contours on the second floor, 5 seconds after door failure, 1.83 m (6 ft) above the floor.}
\label{fig:temp_top_165}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=.7\textwidth]{../Figures/west_50th_baseline_top_220_6ft}
\input{../Scripts/colorbar_std} 
\input{../Scripts/colorbar_metric}
\caption{Temperature contours on the second floor, 60 seconds after door failure, 1.83 m (6 ft) above the floor.}
\label{fig:temp_top_220}
\end{figure}

\section{Tactical Considerations}
In the last fifteen years, there have been fires throughout the country where the change in a flow path has had a significant impact on firefighter and occupant safety. Table~\ref{tab:lodd} lists the NIOSH investigation report number, the outcome, and description of the flow path. It is clear that fires with rapid developing/changing flow paths, whether fire service induced, structure-failure induced, or nature induced are a significant hazard to the fire service. Prediction of the failure mode and failure time of the second floor door is not a reasonable expectation given the large set of unknowns the fire service faced on the fire ground. However, the timing of an interior attack on the structure given the state of fire growth in the rear of the structure identified by a size up, may not have been ideal. Research has shown that applying water from the exterior \cite{madrzykowski2009fire,kerber2009fire} into the structure can have a significant impact on controlling the fire prior to beginning interior operations.

\begin{table}
\centering
\captionof{table}{Flowpath LODD/LODI Incidents}\label{tab:lodd}
\begin{tabular}{l l l}
\toprule[1.5pt]
NIOSH Report $\#$ & $\#$ LODD/LODI & Flow Path Details  \\
\midrule
\multirow{2}{*}{99-F01 \cite{NIOSH:Pettit} }    & \multirow{2}{*}{3 LODD}                                     & \multirow{2}{*}{\parbox{8cm} {From apartment into hallway on 10th floor of high-rise apartment building}} \\
                 &                                                    & \\
99-F21 \cite{NIOSH:Washenitz}         & 2 LODD/2 LODI                          & Basement to 1st floor \\
F2000-04 \cite{NIOSH:Mezzanotte}   & 3 LODD/3 civilian deaths            & 1st floor to 2nd floor \\
F2000-16 \cite{NIOSH:McFall} & 1 LODD/1 civilian death/1 LODI & 2nd floor hallway through 2nd floor apartment \\
\multirow{2}{*}{F2000-23 \cite{NIOSH:McFall2}} & \multirow{2}{*}{1 LODD/2 LODI       }                   & \multirow{2}{*}{\parbox{8cm} {From ground level to 1st floor then to 2nd floor, flow exited through ceiling}} \\
                 &                                                    & \\
F2000-43 \cite{NIOSH:McFall3}         & 1 serious LODI/2 other LODI      & 1st floor to 2nd floor \\
F2004-02  \cite{NIOSH:Berardinelli}  & 1 LODD                                       & 1st floor to basement \\
F2005-02 \cite{NIOSH:Koedam}        & 1 LODD/4 LODI                           & C side to A side of the building \\
F2005-04 \cite{NIOSH:McFall4}         & 1 LODD/9 LODI                           & Basement to 1st floor \\
F2007-09 \cite{NIOSH:Tarley}           & 1 LODD/2 LODI                           & 3 story training burn $-$ flow through all levels \\
F2007-35 \cite{NIOSH:Braddee}       & 4 LODI                                         & 1st floor to 2nd floor \\
F2009-11 \cite{NIOSH:Merinar}         & 2 LODD                                        & C side to A side of the building \\
F2011-13 \cite{NIOSH:Bowyer2}       & 2 LODD                                        & Lower level up stairs and through entry door and garage \\
F2011-31 \cite{NIOSH:Loflin}            & 1 LODD                                        & Fire extended from lower level apartment \\
F2012-28 \cite{NIOSH:Bowyer}        & 1 LODD/1LODI                            & Attic fire extended into closed porch and then into 2nd floor \\
\bottomrule[1.25pt]
\end{tabular}\par
\end{table}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=.80\textwidth]{../Figures/50th_Place_1st_Floor}
\caption {Dimensioned drawing of 1$^{st}$ floor.}
\label{fig:first_floor}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=.80\textwidth]{../Figures/50th_Place_2nd_Floor}
\caption {Dimensioned drawing of 2$^{nd}$ floor.}
\label{fig:second_floor}
\end{figure}
